---
title: "prepare data"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("tidyverse")
library("here")
library("lubridate")
library("xlsx")
```


# Datos a partir de Linaria 

```{r}
avistamientos <- read_csv(here::here("/data/db_linaria/adis_avistamientos.csv")) %>% 
  dplyr::select(-created_at, -updated_at, -observaciones, -the_geom, -accuracy)


visitas <- read_csv(here::here("/data/db_linaria/adis_visitas.csv")) %>% 
  dplyr::select(-created_at, -updated_at, -dicc_viento_id, -dicc_nube_id, -niebla)


# Transectos 
dicc_unidadesmuestreo <- read_csv(here::here("/data/db_linaria/dicc_unidadesmuestreos.csv")) 
dicc_habitat <- read_csv(here::here("/data/db_linaria/dicc_habitats.csv"))

dicc_unidadesmuestreo <- dicc_unidadesmuestreo %>% 
  full_join(dicc_habitat, by=c("dicc_habitat_id"="id")) %>% 
  dplyr::select(-dicc_habitat_id)
  

transectos <- read_csv(here::here("/data/db_linaria/adis_transectos.csv")) %>% 
  dplyr::select(-created_at, -updated_at)

transectos <- transectos %>% inner_join(dicc_unidadesmuestreo, 
                         by = c("dicc_unidadesmuestreo_id" = "id")) %>% 
  dplyr::select(-dicc_unidadesmuestreo_id)

rm(dicc_habitat, dicc_unidadesmuestreo)
# ---------------


# sps 
adis_species <- read_csv(here::here("/data/db_linaria/adis_especies_dispersantes.csv")) %>% 
  dplyr::select(-created_at, -updated_at) 

dicc_species <- read_csv(here::here("/data/db_linaria/dicc_especies.csv")) %>% 
  dplyr::select(-created_at, -updated_at)

adis_species <- adis_species %>% 
  inner_join(dicc_species, 
             by = c("dicc_especie_id" = "id")) %>% 
  dplyr::select(-dicc_especie_id, -imagen, -audio)

#-----------------

# Selecciono los transectos de robledal 

transectos_qp <- transectos %>% 
  filter(habitat == "Robledal") %>% 
  mutate(loc = case_when(
    str_detect(nombre, "Cáñar") ~ "CAN", 
    str_detect(nombre, "Dílar") ~ "DIL",
    str_detect(nombre, "Hornillo") ~ "GEN",
    str_detect(nombre, "Camarate") ~ "CAM"
  )) %>% 
  dplyr::select(-habitat)
write_csv(transectos_qp, here::here("/data/transectos_qp.csv"))

# Selecciono visitas a transectos de robledal 
visitas_qp <- visitas %>% inner_join(transectos_qp, by = c("adis_transecto_id" = "id")) %>% 
  mutate(date = ymd(as.Date(fechai, format="%Y-%m-%d"))) %>% 
  dplyr::select(-adis_transecto_id, -longitud, -fechai, -fechaf) %>% 
  mutate(year = year(date))

# dataframe 
df <- avistamientos %>% 
  inner_join(visitas_qp, by = c("adis_visita_id" = "id")) %>% 
  dplyr::select(-adis_visita_id) %>% 
  inner_join((adis_species %>% dplyr::select(id, nombre_cientifico)), 
             by = c("adis_especies_dispersante_id" = "id")) %>% 
  dplyr::select(-adis_especies_dispersante_id, 
                specie = nombre_cientifico) %>% 
  relocate(id, specie, numero:desplazamiento, loc, date, year)

write_csv(df, here::here("/data/dispersantes_qp.csv"))
```


- Cuando hemos hecho los análisis parece que se han detectado errores: por ejemplo no están todos los transectos; para el arrendajo existen en Camarate varios años sin datos. 

# Datos corregidos enviados por JMBarea 

Either NULL to guess all from the spreadsheet or a character vector containing one entry per column from these options: "skip", "guess", "logical", "numeric", "date", "text" or "list". If exactly one col_type is specified, it will be recycled. The content of a cell in a skipped column is never read and that column will not appear in the data frame output. A list cell loads a column as a list of length 1 vectors, which are typed using the type guessing logic from col_types = NULL, but on a cell-by-cell basis.



```{r}
# d <- readxl::read_excel(here::here("data/Total_ aves_2019_inclusdefinitiva.xlsx")) 

d <- read_delim(here::here("data/Total_ aves_2019_inclusdefinitiva.txt"), delim = "\t",   locale = locale(encoding = 'UTF-8'))

aves_robledal <- d %>% 
  filter(HABITAT == "Robledal") %>% 
  mutate(loc = case_when(
    grepl("Camarate", NOMBRE_TRANSECTO) ~ "CAM",
    grepl("Hornillo", NOMBRE_TRANSECTO) ~ "GEN",
    grepl("Robledal de D", NOMBRE_TRANSECTO) ~ "DIL",
    grepl("Robledal de C", NOMBRE_TRANSECTO) ~ "CAN"
  )) %>% 
  dplyr::select(Ord, PK_VISITAS, FECHA, year = FECHA_A, loc, 
                specie = NOMBRE_CIEN, n= No_IND,  dis = DISTANCIA, desp = DESPLAZAMIENTO) 

write_csv(aves_robledal, here::here("data/dispersantes_qp_new.csv"))
```



